---
title: 
  "Introduction to R Markdown"
author: 
  "Tong Zhu"
date: 
  '27 October 2020'
header-inclues:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{amsthm}  
  - \usepackage{amsthm}  
  - \usepackage{listings}
output: 
  pdf_document:
    fig_width: 6
    fig_height: 6
---

# Abstract

In this file, Latex code, R code and R ouput are written, run/complied and presented as a whole automatically. This is known as _R Markdown_ file. It is particularly useful for writing reports in data science.  

# Introduction

There are a large number of jobs that you could do with _R Markdown_:

  + Compile a single R Markdown document to a report in different formats, such as PDF or HTML.

  + Create notebooks in which you can directly run code chunks interactively.

  + Make slides for presentations (HTML5, LaTeX Beamer, or PowerPoint).

  + Produce dashboards with flexible, interactive, and attractive layouts.

  + Build interactive applications based on Shiny.

  + Write journal articles.

  + Author books of multiple chapters.

  + Generate websites and blogs.

In other words, this is kind of a Latex file. The difference is that it has fewer latex features, but in return, the syntax is easier to handle and you can incorporate R code into it. See the three files that I uploaded on Canvas for more information.  


```{r eval = FALSE}
render("lab3_intro.Rmd")
```

# Methods 

## Math in R Markdown

Usual syntax in Latex works in _R Markdown_. Inline math is marked off with a pair of dollar
signs (`$`), as $\left( \mathbf{X}^{\operatorname{T}}\mathbf{X}\right)^{-1}\mathbf{X}^{\operatorname{T}}\mathbf{y}$. Mathematical displays are marked off with `\[` and `\]`, as in
\[
  f\left(x ; \mu, \sigma^2\right) = \frac{1}{\sqrt{2\pi\sigma^2}} \exp \left( - \frac{\left( x-\mu\right)^2}{2\sigma^2}\right)
\]


## R code

A _R Markdown_ file combines your Latex file with your R script file. 

```{r}
my.mean.func = function(x){
  mean(x)
}

my.mean.func(1:10)
```


## R output

By default R code will be run. But we can adjust things to be included in the pdf file. 

Notice $a$, $b$, $c$ and $d$ are all generated in R. 

$a$: Both the statement and the output are shown in the pdf file. 
```{r}
a = rnorm(10)
a
```

$b$: Only the output is shown in the pdf file, but not the statement
```{r echo = FALSE}
b = rnorm(10)
b
```

$c$: Only the statement is shown in the pdf file, but not output.
```{r results = 'hide'}
c = rnorm(10)
c
```

$d$: Neither the statement nor the output is shown in the pdf file. 
```{r include = FALSE}
d = rnorm(10) 
d
```

## Inline R Code

Of course, sometime you need a number from your data analysis. This can be easily obtained and updated in _R Markdown_ file, e.g. There are `r length(a)` elements in $a$, and the mean of $b$ is `r mean(b)`. 

## Plot 

One of the best features of R markdown is graphs are automatically generated and included in the pdf file. 

```{r}
library(MASS)
data(cats)
plot(Hwt ~ Bwt, data = cats, xlab = "Body weight (kg)", ylab = "Heart weight (g)")
```


There are a bunch of options for adjusting the placement of the plot which
R produces.  

```{r, fig.align="center", out.width="50%", out.height='50%'}
plot(Hwt ~ Bwt, data = cats, xlab = "Body weight (kg)", ylab = "Heart weight (g)")
```

+ The option `fig.align` controls the horizontal **alignment** (left, right,
or center).

+ The options `out.height` and `out.width` let you specify
the desired height or width of the figure. 

## Tables 

The default summary output in R is too ugly to be included in a formal report. Compare the followings:

```{r without knitr}
coefficients(summary(lm(Hwt ~ Bwt, data=cats)))
```


```{r with knitr}
library(knitr) # Only need this the first time!
kable(coefficients(summary(lm(Hwt ~ Bwt, data=cats))))
```

## Computationally intense task

By default, R Markdown will re-run all of your code every time you render/knit your
document.  If some of your code is slow, this can add up to a lot of time.  You
can, however, ask R Markdown to keep track of whether a chunk of code has
changed, and only re-run it if it has.  This is called **caching** the chunk.

```{r, cache=TRUE}
data.sim.vec = rnorm(1e6)
x.vec = sort(data.sim.vec)
y = max(x.vec)
```


# Results, conclusions and recommendations 

In addition to executing R code chunks, it can also execute chunks in a variety of other languages that are widely used in data science. Some of the available language engines include:

+ Python

+ SQL

+ Bash

+ Rcpp

+ Stan

+ JavaScript

+ CSS

See the book ["_R Markdown: The Definitive Guide_"][Yihui Xie, J. J. Allaire, Garrett Grolemund] if you are interested. 


# Conclusion

Use it! 



[Yihui Xie, J. J. Allaire, Garrett Grolemund]: https://bookdown.org/yihui/rmarkdown/


